---
title: "clustering_assessment"
author: "Dmitrii Belousov"
date: "2025-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      root.dir = "~/GVHD_project/notebooks/EDA/")
library(tidyverse)
library(ggplot2)
library(ggridges)
library(GGally)
library(ggpubr)
library(gridExtra)
library(grid)
library(cluster)
library(pheatmap)
library(caret)
library(survival)
library(survminer)
library(NMF)
library(fpc)
library(factoextra)
library(rstatix)
library(corrplot)

set.seed(42)
```

# Data Import & Cleansing

```{r}
data <- read_csv("~/GVHD_project/data/data_wide_preproc.csv")
data <- data %>% 
  mutate(Event = ifelse(is.na(cGVHD_time), 0, 1),
         Time_CFS = ifelse(is.na(cGVHD_time), Time_OS, cGVHD_time))

row.names(data) <- data %>% 
  mutate(ID_new = paste0(round(runif(231, 0, 1000), 0), Collection_time, ID)) %>% 
  pull(ID_new)
```

# Clustering Raw Matrix

## Number of clusters determination

```{r}
hclust_func <- function(data, k) {
  data %>% 
    dist(method = "euclidean") %>% 
    hclust(method = "ward.D") %>% 
    cutree(k = k) -> labels
  
  return(list(clusters = labels))
}

data_subset <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>% 
  select(where(is.numeric), -c(Post_cGVHD_time, ID, Time_OS, cGVHD_time, 
                               Collection_Day, Time_CFS, Event)) %>% 
  #select(contains("226")) %>% 
  log1p() %>% 
  scale() %>% 
  as.data.frame()

rownames(data_subset) <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>%
  mutate(ID_new = paste0(Collection_time, ID)) %>% 
  pull(ID_new)

fviz_nbclust(data_subset, FUNcluster = hclust_func, method = "silhouette")
```

## Clustering

```{r, fig.height=5, fig.width=10}

res <- data_subset %>% 
  dist(method = "euclidean") %>% 
  hclust(method = "ward.D", ) 

res %>% 
  factoextra::fviz_dend(cex = 0.2, 
            k = 3, 
            k_colors = "jco")
```

```{r, fig.height=10, fig.width=25}
labels <- res %>% 
  cutree(k = 3)

data_subset_sorted <- data_subset %>% 
  mutate(labels = labels) %>% 
  arrange(labels)

labels <- data_subset_sorted %>% 
  pull(labels) %>% 
  as.data.frame()

data_subset_sorted <-data_subset_sorted %>% 
  select(everything(), -labels)

rownames(labels) <- rownames(data_subset_sorted)
labels$. <- as.factor(labels$.)

data_subset_sorted %>% 
  mutate(across(everything(), ~ pmax(pmin(., 2), 0))) %>% 
  pheatmap::pheatmap(cellwidth = 8, 
                     cellheight = 8, 
                     cluster_rows = FALSE,
                     annotation_row = labels,
                     border_color = NA)

```

## Assessment
### Sillhouette

```{r}
sil <- silhouette(res %>% cutree(k = 3), 
                  data_subset %>% dist(method = "euclidean"))
fviz_silhouette(sil) +
  theme_bw()
```

### Jaccard Stability

```{r}
Hclustering <- function(data_matrix, k = 4) {
  clusters <- data_matrix %>% 
    dist(method = "euclidean") %>% 
    hclust(method = "ward.D") %>% 
    cutree(k = k)
  
  if (!is.factor(clusters)) {
    clusters <- factor(clusters)
  }
  
  levels(clusters) <- as.character(0:k-1)

  clist <- lapply(levels(clusters), function(cluster) {clusters == cluster})
  names(clist) <- levels(clusters)

  output = list(
    result = clusters,
    nc = n_distinct(clusters),
    clusterlist = clist,
    partition = rep(1, length(clusters)),
    clustermethod = "WARD"
  )
  
  return(output)
}
```

```{r}
data_subset_mtx <- as.matrix(data_subset)
```

```{r}
cpx <- clusterboot(data_subset_mtx, 
            B = 100, k = 4,
            clustermethod = Hclustering, 
            bootmethod = "boot", 
            multipleboot = TRUE)
```


```{r}
print(cpx)
```

## Survival

```{r}
ds1 <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>% 
  mutate(ID_new = paste0(Collection_time, ID)) %>% 
  select(Time_CFS, Event, ID_new)
  

ds2 <- data_subset_sorted %>% 
  mutate(ID_new = rownames(data_subset_sorted), 
         labels = labels$.)

merged <- merge(ds1, ds2, by.x="ID_new", by.y="ID_new", all=T)

surv_object <- Surv(time = as.numeric(merged$Time_CFS), event = merged$Event)

fit <- survfit(surv_object ~ `labels`, data = merged)

ggsurvplot(fit, 
           data = merged,
           title = "Cluster survival curves",
           conf.int	= FALSE,        
           risk.table = TRUE,
           tables.height = 0.3,
           pval = TRUE,
           pval.coord = c(650, 0.7))         
```

## Cluster characteristics

```{r}
stats <- list()

colnames(labels) <- c("consensus")

data_subset_cc <- cbind(labels, data_subset) %>% 
  filter(consensus != 7) %>% 
  mutate(consensus = as.factor(consensus))

for (i in c(1:n_distinct(data_subset_cc$consensus))) {
  data_internal <- data_subset_cc %>% 
    mutate(across(where(is.numeric), scale)) %>% 
    mutate(test_col = ifelse(consensus == i, i, "other")) %>% 
    select(everything(), -c(consensus))
  
  MD <- data_internal %>% 
    summarise(across(where(is.numeric), 
                       ~ t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate1 - 
                        t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate2))# %>% log2()
  
  Ps <- data_internal %>% 
    summarise(across(where(is.numeric), ~ t_test(.x ~ test_col, data = data_internal, 
                                               detailed = TRUE)$p))
  p.adj <- p.adjust(Ps, method = "BH")
  
  stats[[paste(i, "vs. all")]] <- as.data.frame(t(rbind(MD, Ps, p.adj)))
  colnames(stats[[paste(i, "vs. all")]]) <- c("MeanDiff", "P", "Padj")
}
```

```{r, fig.width=10, fig.height = 4}
library(cowplot)

plot_MD <- function(data){
  d <- data %>% 
    mutate(label = rownames(data)) %>% 
    filter(Padj < 0.05) %>% 
    arrange(desc(abs(MeanDiff))) %>% 
    slice_head(n = 10) %>% 
    arrange(desc(MeanDiff)) %>% 
    mutate(position = seq(10, 1)) 
  
  min <- min(d$MeanDiff)
  max <- max(d$MeanDiff)
  
  p <- ggplot(d, aes(x = MeanDiff, y = position)) +          
      geom_text(aes(label = label),       
                vjust = -0.5,              
                hjust = 0,             
                color = "black") +  
      xlim(min, max + 0.5) +
      ylim(.5, 10.5) +
      theme_bw()
  
  return(p)
}

plot_grid(plotlist = lapply(stats, plot_MD), ncol=3)
```


# Clustering Correlation matrix

```{r}
collection_time_point_palette <- c(
  "Never" = "#00AA96", 
  "Pre" = "#F88F70",
  "Post" = "#F56376"
)
```

## Correlation analysis

```{r, fig.height=10, fig.width=10}
data_subset <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>% 
  select(where(is.numeric), -c(Post_cGVHD_time, ID, Time_OS, cGVHD_time, Collection_Day, Event, Time_CFS)) %>% 
  #select(contains("226")) %>% 
  log1p() %>% 
  scale() %>% 
  t() %>% 
  as.data.frame() 

colnames(data_subset) <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>%
  mutate(ID_new = paste0(Collection_time, ID)) %>% 
  pull(ID_new)

annotation_row_vec <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>%
  pull(Collection_time) %>% 
  as.data.frame()

rownames(annotation_row_vec) <- colnames(data_subset)

corr_data <- data_subset %>% 
  as.data.frame() %>% 
  psych::corr.test(method = "spearman")

corr_data$r %>% 
  pheatmap::pheatmap(cellwidth = 8, 
                     cellheight = 8, 
                     border_color = NA, 
                     annotation_row = annotation_row_vec, 
                     annotation_colors = list(. = collection_time_point_palette))
```

## Number of clusters determination

```{r}
fviz_nbclust(corr_data$r, FUNcluster = hclust_func, method = "silhouette")
```

## Clustering

```{r}
cor_res <- corr_data$r %>% 
  dist(method = "euclidean") %>% 
  hclust(method = "ward.D") 

cor_res %>% 
  factoextra::fviz_dend(cex = 0.2, 
            k = 4, 
            k_colors = "jco")

labels <- cor_res %>% 
  cutree(k = 4)
```

```{r, fig.height=10, fig.width=25}
data_subset_sorted <- data_subset %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(labels = labels) %>% 
  arrange(labels)

labels <- data_subset_sorted %>% 
  pull(labels) %>% 
  as.data.frame()

data_subset_sorted <-data_subset_sorted %>% 
  select(everything(), -labels)

rownames(labels) <- rownames(data_subset_sorted)
labels$. <- as.factor(labels$.)

data_subset_sorted %>% 
  mutate(across(everything(), ~ pmax(pmin(., 2), 0))) %>% 
  pheatmap::pheatmap(cellwidth = 8, 
                     cellheight = 8, 
                     cluster_rows = FALSE,
                     annotation_row = labels,
                     border_color = NA)

```

## Assessment
### Sillhouette

```{r}
sil <- silhouette(cor_res %>% cutree(k = 4), 
                  corr_data$r %>% dist(method = "euclidean"))
fviz_silhouette(sil) +
  theme_bw()
```

### Jaccard Similarity

```{r}
r_mtx <- as.matrix(corr_data$r)

cpx <- clusterboot(r_mtx, 
            B = 100, k = 4,
            clustermethod = Hclustering, 
            bootmethod = "boot", 
            multipleboot = FALSE)
```

```{r}
print(cpx)
```

## Survival

```{r}
ds1 <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>% 
  mutate(ID_new = paste0(Collection_time, ID)) %>% 
  select(Time_CFS, Event, ID_new, Collection_time)
  

ds2 <- data_subset_sorted %>% 
  mutate(ID_new = rownames(data_subset_sorted), 
         labels = labels$.) 

merged <- merge(ds1, ds2, by.x="ID_new", by.y="ID_new", all=T)

surv_object <- Surv(time = as.numeric(merged$Time_CFS), event = merged$Event)

fit <- survfit(surv_object ~ `labels`, data = merged)

ggsurvplot(fit, 
           data = merged,
           title = "Cluster survival curves",
           conf.int	= FALSE,        
           risk.table = TRUE,
           tables.height = 0.3,
           pval = TRUE,
           pval.coord = c(650, 0.7)) 
```


## Cluster characteristics

```{r}
stats <- list()

colnames(labels) <- c("consensus")

data_subset <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>% 
  select(where(is.numeric), -c(Post_cGVHD_time, ID, Time_OS, cGVHD_time, 
                               Collection_Day, Time_CFS, Event)) %>% 
  #select(contains("226")) %>% 
  log1p() %>% 
  scale() %>% 
  as.data.frame()

rownames(data_subset) <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>%
  mutate(ID_new = paste0(Collection_time, ID)) %>% 
  pull(ID_new)

data_subset_cc <- cbind(labels, data_subset) %>% 
  filter(consensus != 7) %>% 
  mutate(consensus = as.factor(consensus))

for (i in c(1:n_distinct(data_subset_cc$consensus))) {
  data_internal <- data_subset_cc %>% 
    mutate(across(where(is.numeric), scale)) %>% 
    mutate(test_col = ifelse(consensus == i, i, "other")) %>% 
    select(everything(), -c(consensus))
  
  MD <- data_internal %>% 
    summarise(across(where(is.numeric), 
                       ~ t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate1 - 
                        t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate2))# %>% log2()
  
  Ps <- data_internal %>% 
    summarise(across(where(is.numeric), ~ t_test(.x ~ test_col, data = data_internal, 
                                               detailed = TRUE)$p))
  p.adj <- p.adjust(Ps, method = "BH")
  
  stats[[paste(i, "vs. all")]] <- as.data.frame(t(rbind(MD, Ps, p.adj)))
  colnames(stats[[paste(i, "vs. all")]]) <- c("MeanDiff", "P", "Padj")
}
```

```{r, fig.width=10, fig.height = 10}
library(cowplot)

plot_MD <- function(data){
  d <- data %>% 
    mutate(label = rownames(data)) %>% 
    filter(Padj < 0.05) %>% 
    arrange(desc(abs(MeanDiff))) %>% 
    slice_head(n = 10) %>% 
    arrange(desc(MeanDiff)) 
  
  if (nrow(d) > 0) {
    d$position = seq(nrow(d), 1)
  } else {
    return(
      ggplot() +
        annotate("text", 
                 x = 3, y = 6, label = "No Sign. Features", 
                 size = 5, color = "red") +
        theme_bw()
    )
  }
  
  min <- min(d$MeanDiff)
  max <- max(d$MeanDiff)
  
  p <- ggplot(d, aes(x = MeanDiff, y = position)) +          
      geom_text(aes(label = label),       
                vjust = -0.5,              
                hjust = 0,             
                color = "black") +  
      xlim(min, max + 0.5) +
      ylim(.5, 10.5) +
      theme_bw()
  
  return(p)
}

plot_grid(plotlist = lapply(stats, plot_MD))
```

# Clustering Embedding

## PCA

```{r}
data_subset <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>% 
  select(where(is.numeric), -c(Post_cGVHD_time, ID, Time_OS, cGVHD_time, 
                               Collection_Day, Time_CFS, Event)) %>% 
  #select(contains("226")) %>% 
  log1p() %>% 
  scale() %>% 
  as.data.frame()

rownames(data_subset) <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>%
  mutate(ID_new = paste0(Collection_time, ID)) %>% 
  pull(ID_new)
```

### Picking appropriate factorization

```{r}
data_subset %>% 
  prcomp() %>% 
  fviz_eig(addlabels = TRUE)
 
```

### Factorization & Visualization

```{r}
data_subset %>% 
  prcomp(rank. = 10) -> res.pca

fviz_pca_ind(res.pca, 
             label="none", 
             habillage=gsub("[0-9]", "", rownames(data_subset)),
             addEllipses=TRUE, 
             ellipse.level=0.95, 
             palette = collection_time_point_palette, 
             title = "Cohort at day 90")
```

```{r}
fviz_contrib(res.pca, choice="var", axes = 1, top = 10)
fviz_contrib(res.pca, choice="var", axes = 2, top = 10)
```


### Number of clusters determination

```{r}
fviz_nbclust(res.pca$x, FUNcluster = kmeans, method = "silhouette")
```

### Clustering

```{r}
res <- kmeans(res.pca$x, centers = 4, nstart = 100)
data_subset$labels <- as.factor(res$cluster)

fviz_pca_ind(res.pca, 
             label="none", 
             habillage=data_subset$labels,
             addEllipses=TRUE, 
             ellipse.level=0.95, 
             palette = "Dark2", 
             title = "Cohort at day 90 Clustering")
```

```{r, fig.height=10, fig.width=25}
data_subset_sorted <- data_subset %>% 
  arrange(labels)

labels <- data_subset_sorted %>% 
  pull(labels) %>% 
  as.data.frame()

data_subset_sorted <-data_subset_sorted %>% 
  select(everything(), -labels)

rownames(labels) <- rownames(data_subset_sorted)
labels$. <- as.factor(labels$.)

data_subset_sorted %>% 
  mutate(across(everything(), ~ pmax(pmin(., 2), 0))) %>% 
  pheatmap::pheatmap(cellwidth = 8, 
                     cellheight = 8, 
                     cluster_rows = FALSE,
                     annotation_row = labels,
                     border_color = NA)
```

### Assessment
#### Sillhouette

```{r}
sil <- silhouette(as.numeric(data_subset$labels), 
                  res.pca$x %>% dist(method = "euclidean"))
fviz_silhouette(sil) +
  theme_bw()
```

#### Jaccard Similarity

```{r}
KMEANSclustering <- function(data_matrix, k = 4) {
  res <- data_matrix %>% 
    kmeans(centers = k, nstart = 100)
  
  clusters <- res$cluster
  
  if (!is.factor(clusters)) {
    clusters <- factor(clusters)
  }
  
  levels(clusters) <- as.character(0:k-1)

  clist <- lapply(levels(clusters), function(cluster) {clusters == cluster})
  names(clist) <- levels(clusters)

  output = list(
    result = clusters,
    nc = n_distinct(clusters),
    clusterlist = clist,
    partition = rep(1, length(clusters)),
    clustermethod = "KMEANS"
  )
  
  return(output)
}
```



```{r}
cpx <- clusterboot(as.matrix(res.pca$x), 
            B = 100, k = 4,
            clustermethod = KMEANSclustering, 
            bootmethod = "boot", 
            multipleboot = FALSE)
```
```{r}
print(cpx)
```


### Survival

```{r}
ds1 <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>% 
  mutate(ID_new = paste0(Collection_time, ID)) %>% 
  select(Time_CFS, Event, ID_new)
  

ds2 <- data_subset_sorted %>% 
  mutate(ID_new = rownames(data_subset_sorted), 
         labels = labels$.)

merged <- merge(ds1, ds2, by.x="ID_new", by.y="ID_new", all=T)

surv_object <- Surv(time = as.numeric(merged$Time_CFS), event = merged$Event)

fit <- survfit(surv_object ~ `labels`, data = merged)

ggsurvplot(fit, 
           data = merged,
           title = "Cluster survival curves",
           conf.int	= FALSE,        
           risk.table = TRUE,
           tables.height = 0.3,
           pval = TRUE,
           pval.coord = c(650, 0.7)) 
```


### Cluster characteristics

```{r}
stats <- list()

colnames(labels) <- c("consensus")

data_subset <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>% 
  select(where(is.numeric), -c(Post_cGVHD_time, ID, Time_OS, cGVHD_time, 
                               Collection_Day, Time_CFS, Event)) %>% 
  #select(contains("226")) %>% 
  log1p() %>% 
  scale() %>% 
  as.data.frame()

rownames(data_subset) <- data %>% 
  filter(is.na(cGVHD_day)) %>% 
  filter(!is.na(Collection_time) & Collection_Day %in% c("90") & Collection_time != "Post") %>%
  mutate(ID_new = paste0(Collection_time, ID)) %>% 
  pull(ID_new)

data_subset_cc <- cbind(labels, data_subset) %>% 
  mutate(consensus = as.factor(consensus))

for (i in c(1:n_distinct(data_subset_cc$consensus))) {
  data_internal <- data_subset_cc %>% 
    mutate(across(where(is.numeric), scale)) %>% 
    mutate(test_col = ifelse(consensus == i, i, "other")) %>% 
    select(everything(), -c(consensus))
  
  MD <- data_internal %>% 
    summarise(across(where(is.numeric), 
                       ~ t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate1 - 
                        t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate2))# %>% log2()
  
  Ps <- data_internal %>% 
    summarise(across(where(is.numeric), ~ t_test(.x ~ test_col, data = data_internal, 
                                               detailed = TRUE)$p))
  p.adj <- p.adjust(Ps, method = "BH")
  
  stats[[paste(i, "vs. all")]] <- as.data.frame(t(rbind(MD, Ps, p.adj)))
  colnames(stats[[paste(i, "vs. all")]]) <- c("MeanDiff", "P", "Padj")
}
```

```{r, fig.width=10, fig.height = 10}
library(cowplot)

plot_MD <- function(data){
  d <- data %>% 
    mutate(label = rownames(data)) %>% 
    filter(Padj < 0.05) %>% 
    arrange(desc(abs(MeanDiff))) %>% 
    slice_head(n = 10) %>% 
    arrange(desc(MeanDiff)) 
  
  if (nrow(d) > 0) {
    d$position = seq(nrow(d), 1)
  } else {
    return(
      ggplot() +
        annotate("text", 
                 x = 3, y = 6, label = "No Sign. Features", 
                 size = 5, color = "red") +
        theme_bw()
    )
  }
  
  min <- min(d$MeanDiff)
  max <- max(d$MeanDiff)
  
  p <- ggplot(d, aes(x = MeanDiff, y = position)) +          
      geom_text(aes(label = label),       
                vjust = -0.5,              
                hjust = 0,             
                color = "black") +  
      xlim(min, max + 0.5) +
      ylim(.5, 10.5) +
      theme_bw()
  
  return(p)
}

plot_grid(plotlist = lapply(stats, plot_MD))
```

## NMF (scaled)
### Picking appropriate factorization rank

```{r}
normalize <- function(x, na.rm = TRUE) {
    return((x - min(x)) /(max(x) - min(x)))
}

data %>% 
  filter(is.na(cGVHD_day) 
         & !is.na(Collection_time) 
         & Collection_Day %in% c("90") 
         & Collection_time != "Post") %>% 
  select(everything(), -c(ID, Time_OS, Time_CFS, cGVHD_time, cGVHD_day, 
                          Collection_Day, Source, cGVHD_time_discrete, 
                          Collection_time, Post_cGVHD_time)) %>% 
  mutate(across(-Event, log1p), 
         across(-Event, normalize), 
         Event = as.factor(Event)) -> data_filtered

event <- data_filtered %>% 
  pull(Event)

data_filtered <- data_filtered %>% 
  select(everything(), -Event) %>% 
  t() %>% 
  as.data.frame()

data_filtered_random <- randomize(data_filtered)
```

```{r}
estim.r.random <- nmf(data_filtered_random, 
                      2:15, 
                      nrun=10, 
                      seed=42)

estim.r <- nmf(data_filtered, 
               2:15, 
               nrun=10, 
               seed=42)
```

```{r, fig.height=7, fig.width=12}
plot(estim.r, estim.r.random)
```

```{r}
estim.r <- nmf(data_filtered, 
               5, 
               nrun = 100, 
               seed = 42, 
               .options = "t", 
               .pbackend = 6)

plot(estim.r)
```

### Cluster Stability 

```{r}
NMFclustering <- function(data_matrix, k = 5) {
  res <- nmf(t(data_matrix), k, nrun=100, seed=42)
  clusters <- predict(res)
  if (!is.factor(clusters)) {
    clusters <- factor(clusters)
  }
  
  levels(clusters) <- as.character(0:k-1)

  clist <- lapply(levels(clusters), function(cluster) {clusters == cluster})
  names(clist) <- levels(clusters)

  output = list(
    result = clusters,
    nc = n_distinct(clusters),
    clusterlist = clist,
    partition = rep(1, length(clusters)), 
    clustermethod = "NMF"
  )
  
  return(output)
}

```

```{r}
mtx <- as.matrix(data_filtered)
```

```{r}
cpx <- clusterboot(t(mtx), 
            B = 10, 
            clustermethod = NMFclustering, 
            bootmethod = "boot", 
            multipleboot = TRUE)

print(cpx)
```

```{r fig.height=5, fig.width=8}
res <- nmf(data_filtered, 5, nrun=100, seed=42)

layout(cbind(1,2))
basismap(res)
coefmap(res)
```

```{r}
consensusmap(res, labCol=NA, labRow=NA)
```

```{r}
event <- as.data.frame(event)
rownames(event) <- paste0("V", rownames(event))

cc <- as.data.frame(predict(res))
colnames(cc) <- c("consensus")
```

```{r, fig.height=7, fig.width=10}
library(RColorBrewer)

event_colors <- setNames(brewer.pal(length(unique(event$event)), "Set2"), unique(event$event))
cc_colors <- setNames(brewer.pal(length(unique(cc$consensus)), "Pastel1"), unique(cc$consensus))

annotation_colors <- list(event = event_colors, consensus = cc_colors)

order <- cc %>% 
  arrange(consensus) %>% 
  rownames()

pheatmap::pheatmap(res@consensus[order, order], 
                     cellwidth = 6, cellheight = 6, 
                     cluster_rows = F, cluster_cols = F,
                     annotation_col = cbind(event, cc),
                     annotation_colors = annotation_colors,
                     border_color = NA)
```

### Cluster charcterization

```{r}
W <- res@fit@W
colnames(W) <- c(1:5)
top_features_all <- lapply(1:ncol(W), function(cluster) {
  sort(W[, cluster], decreasing = TRUE)[1:10]
})

names(top_features_all) <- paste0("Cluster_", 1:ncol(W))
print(top_features_all)

```

```{r, fig.height=6}
top_feature_names <- unique(unlist(lapply(top_features_all, names)))
W_top <- W[top_feature_names, ] 

pheatmap::pheatmap(t(W_top), 
                   cellwidth = 10, 
                   cellheight = 8, 
                   cluster_rows = F,
                   show_rownames = T,
                   cluster_cols = T, 
                   border_color = NA, 
                   scale = "row")
```

```{r}
stats <- list()

colnames(labels) <- c("consensus")

data_filtered_full <- data %>% 
  filter(is.na(cGVHD_day) 
         & !is.na(Collection_time) 
         & Collection_Day %in% c("90") 
         & Collection_time != "Post") %>% 
  as.data.frame()

data_subset_cc <- cbind(cc, data_filtered_full) %>% 
  #filter(consensus != 7) %>% 
  mutate(consensus = as.factor(consensus))

for (i in c(1:n_distinct(data_subset_cc$consensus))) {
  data_internal <- data_subset_cc %>% 
    mutate(test_col = ifelse(consensus == i, i, "other")) %>% 
    select(everything(), -c(ID, Time_OS, Time_CFS, cGVHD_time, cGVHD_day, 
                          Collection_Day, Source, cGVHD_time_discrete, 
                          Collection_time, Post_cGVHD_time, Event, consensus))
  
  log2FC <- data_internal %>% 
    summarise(across(where(is.numeric), 
                       ~ t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate1 / 
                        t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate2)) %>% log2()
  
  Ps <- data_internal %>% 
    summarise(across(where(is.numeric), ~ t_test(.x ~ test_col, data = data_internal, 
                                               detailed = TRUE)$p))
  
  p.adj <- p.adjust(Ps, method = "BH")
  
  stats[[paste(i, "vs. all")]] <- as.data.frame(t(rbind(log2FC, Ps, p.adj)))
  colnames(stats[[paste(i, "vs. all")]]) <- c("LFC", "P", "Padj")
}
```

```{r, fig.width=10, fig.height = 10}
library(cowplot)

plot_LFC <- function(data){
  d <- data %>% 
    mutate(label = rownames(data)) %>% 
    filter(Padj < 0.05) %>% 
    arrange(desc(abs(LFC))) %>% 
    slice_head(n = 10) %>% 
    arrange(desc(LFC)) 
  
  if (nrow(d) > 0) {
    d$position = seq(nrow(d), 1)
  } else {
    return(
      ggplot() +
        annotate("text", 
                 x = 3, y = 6, label = "No Sign. Features", 
                 size = 5, color = "red") +
        theme_bw()
    )
  }
  
  min <- min(d$LFC)
  max <- max(d$LFC)
  
  p <- ggplot(d, aes(x = LFC, y = position)) +          
      geom_text(aes(label = label),       
                vjust = -0.5,              
                hjust = 0,             
                color = "black") +  
      xlim(min, max + 1) +
      ylim(.5, 10.5) +
      theme_bw()
  
  return(p)
}

plot_grid(plotlist = lapply(stats, plot_LFC))
```

## NMF (log)
### Picking appropriate factorization rank

```{r}
data %>% 
  filter(is.na(cGVHD_day) 
         & !is.na(Collection_time) 
         & Collection_Day %in% c("90") 
         & Collection_time != "Post") %>% 
  select(everything(), -c(ID, Time_OS, Time_CFS, cGVHD_time, cGVHD_day, 
                          Collection_Day, Source, cGVHD_time_discrete, 
                          Collection_time, Post_cGVHD_time)) %>% 
  mutate(across(-Event, log1p), 
         # across(-Event, normalize), 
         Event = as.factor(Event)) -> data_filtered

event <- data_filtered %>% 
  pull(Event)

data_filtered <- data_filtered %>% 
  select(everything(), -Event) %>% 
  t() %>% 
  as.data.frame()

data_filtered_random <- randomize(data_filtered)
```

```{r}
estim.r.random <- nmf(data_filtered_random, 
                      2:15, 
                      nrun=10, 
                      seed=42)

estim.r <- nmf(data_filtered, 
               2:15, 
               nrun=10, 
               seed=42)
```

```{r, fig.height=7, fig.width=12}
plot(estim.r, estim.r.random)
```

```{r}
estim.r <- nmf(data_filtered, 
               7, 
               nrun = 100, 
               seed = 42, 
               .options = "t", 
               .pbackend = 6)

plot(estim.r)
```

### Cluster Stability 

```{r}
mtx <- as.matrix(data_filtered)
```

```{r}
cpx <- clusterboot(t(mtx), 
            B = 10, k = 7,
            clustermethod = NMFclustering, 
            bootmethod = "boot", 
            multipleboot = TRUE)

print(cpx)
```

```{r fig.height=5, fig.width=8}
res <- nmf(data_filtered, 7, nrun=100, seed=42)

layout(cbind(1,2))
basismap(res)
coefmap(res)
```

```{r}
consensusmap(res, labCol=NA, labRow=NA)
```

```{r}
event <- as.data.frame(event)
rownames(event) <- paste0("V", rownames(event))

cc <- as.data.frame(predict(res))
colnames(cc) <- c("consensus")
```

```{r, fig.height=7, fig.width=10}
library(RColorBrewer)

event_colors <- setNames(brewer.pal(length(unique(event$event)), "Set2"), unique(event$event))
cc_colors <- setNames(brewer.pal(length(unique(cc$consensus)), "Pastel1"), unique(cc$consensus))

annotation_colors <- list(event = event_colors, consensus = cc_colors)

order <- cc %>% 
  arrange(consensus) %>% 
  rownames()

pheatmap::pheatmap(res@consensus[order, order], 
                     cellwidth = 6, cellheight = 6, 
                     cluster_rows = F, cluster_cols = F,
                     annotation_col = cbind(event, cc),
                     annotation_colors = annotation_colors,
                     border_color = NA)
```

### Cluster charcterization

```{r}
W <- res@fit@W
colnames(W) <- c(1:7)
top_features_all <- lapply(1:ncol(W), function(cluster) {
  sort(W[, cluster], decreasing = TRUE)[1:10]
})

names(top_features_all) <- paste0("Cluster_", 1:ncol(W))
print(top_features_all)

```

```{r}
W %>% write.csv("~/GVHD_project/data/W_nmf.csv")
```

```{r, fig.height=6}
top_feature_names <- unique(unlist(lapply(top_features_all, names)))
W_top <- W[top_feature_names, ] 

W_top %>% write.csv("~/GVHD_project/data/W_top_nmf.csv")

pheatmap::pheatmap(t(W_top), 
                   cellwidth = 10, 
                   cellheight = 8, 
                   cluster_rows = F,
                   show_rownames = T,
                   cluster_cols = T, 
                   border_color = NA, 
                   scale = "row")
```

```{r}
stats <- list()

colnames(labels) <- c("consensus")

data_filtered_full <- data %>% 
  filter(is.na(cGVHD_day) 
         & !is.na(Collection_time) 
         & Collection_Day %in% c("90") 
         & Collection_time != "Post") %>% 
  as.data.frame()

data_subset_cc <- cbind(cc, data_filtered_full) %>% 
  filter(consensus != 7) %>% 
  mutate(consensus = as.factor(consensus))

for (i in c(1:n_distinct(data_subset_cc$consensus))) {
  data_internal <- data_subset_cc %>% 
    mutate(test_col = ifelse(consensus == i, i, "other")) %>% 
    select(everything(), -c(ID, Time_OS, Time_CFS, cGVHD_time, cGVHD_day, 
                          Collection_Day, Source, cGVHD_time_discrete, 
                          Collection_time, Post_cGVHD_time, Event, consensus))
  
  log2FC <- data_internal %>% 
    summarise(across(where(is.numeric), 
                       ~ 1 + (t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate1 / 
                        t_test(.x ~ test_col, 
                                data = data_internal, 
                                detailed = TRUE)$estimate2))) %>% log2()
  
  Ps <- data_internal %>% 
    summarise(across(where(is.numeric), ~ t_test(.x ~ test_col, data = data_internal, 
                                               detailed = TRUE)$p))
  
  p.adj <- p.adjust(Ps, method = "BH")
  
  stats[[paste(i, "vs. all")]] <- as.data.frame(t(rbind(log2FC, Ps, p.adj)))
  colnames(stats[[paste(i, "vs. all")]]) <- c("LFC", "P", "Padj")
}
```

```{r, fig.width=10, fig.height = 10}
library(cowplot)

plot_LFC <- function(data){
  d <- data %>% 
    mutate(label = rownames(data)) %>% 
    filter(Padj < 0.05) %>% 
    arrange(desc(abs(LFC))) %>% 
    slice_head(n = 10) %>% 
    arrange(desc(LFC)) 
  
  if (nrow(d) > 0) {
    d$position = seq(nrow(d), 1)
  } else {
    return(
      ggplot() +
        annotate("text", 
                 x = 3, y = 6, label = "No Sign. Features", 
                 size = 5, color = "red") +
        theme_bw()
    )
  }
  
  min <- min(d$LFC)
  max <- max(d$LFC)
  
  p <- ggplot(d, aes(x = LFC, y = position)) +          
      geom_text(aes(label = label),       
                vjust = -0.5,              
                hjust = 0,             
                color = "black") +  
      xlim(min, max + 1) +
      ylim(.5, 10.5) +
      theme_bw()
  
  return(p)
}

plot_grid(plotlist = lapply(stats, plot_LFC))
```

### Survival

```{r}
data_filtered_t <- data_filtered %>% 
  t() %>% 
  as.data.frame()
```

```{r, fig.height=5}
ds_cfs <- data %>% 
  filter(is.na(cGVHD_day) 
         & !is.na(Collection_time) 
         & Collection_Day %in% c("90") 
         & Collection_time != "Post") %>% 
  select(c(ID, Time_CFS, Event)) %>% 
  as.data.frame()

rownames(ds_cfs) <- rownames(data_filtered_t)

ds_cfs_cc <- cbind(ds_cfs[rownames(cc),], cc) %>% filter(consensus != 7)

#ds_cfs_cc = ds_cfs_cc %>% filter(Event == 1) %>% filter(!(consensus %in% c(4, 6)))

surv_object <- Surv(time = as.numeric(ds_cfs_cc$Time_CFS), event = ds_cfs_cc$Event)

fit <- survfit(surv_object ~ consensus, data = ds_cfs_cc)

ggsurvplot(fit, 
           data = ds_cfs_cc,
           title = "Cluster survival curves",
           conf.int	= FALSE,        
           risk.table = TRUE,
           tables.height = 0.375,
           pval = TRUE,
           #xlim = c(0, 365),
           pval.coord = c(300, 0.99)) 
```

# Feature Selection + Clustering









































