---
title: "Preliminary_EDA"
author: "Dmitrii Belousov"
date: "2024-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggridges)
library(GGally)
library(ggpubr)
library(gridExtra)
library(grid)
```

```{r}
# install.packages("ggridges")
# install.packages("GGally")
# install.packages("ggpubr")
# install.packages("gridExtra")
# install.packages("grid")
renv::snapshot(type = "all")
```
```{r}

palette <- c(
 "4_TFH" = "#3B0A45",  # Deep plum, sophisticated and rich
  "4_TH1" = "#1A3D6B",  # Muted navy blue, professional yet distinct
  "4_TH17" = "#6D9A3B", # Muted olive green for calm, intellectual tone
  "4_Th17TO1" = "#D66F41", # Soft burnt orange, elegant warmth
  "4_TH2" = "#97B498",  # Subtle sage green, soft and approachable
  "4_TH22" = "#9B3D53"  # S
)
```


# Data Import

```{r}
data <- read_csv("../../data/data_wide.csv")
data %>% head()
```
```{r}
data %>% glimpse()
```

# let's look inside

```{r}

palette = c("#3B0A45", "#1A3D6B", "#6D9A3B", "#D66F41", "#97B498", 
            "#9B3D53", "#607DAB", "#B94553", "#0063AB", "#007670", 
            "#D57184")

breaks = c(0.005, 0.05, 0.5, 5, 5, 50, 500)

ridge_plot <- function(data, 
                       palette, 
                       groups = NULL,
                       log_scale = TRUE, 
                       log_breaks = c(0.005, 0.05, 0.5, 5, 5, 50, 500), 
                       drop_tick_labs = FALSE) {
  
  data$FillGroup <- if (is.null(groups)) data$Population else data[[groups]]
  
  plot <- ggplot(data, aes(x = Count, 
                           y = Population, 
                           fill = FillGroup)) +
    geom_density_ridges(rel_min_height = 0.01, alpha = 0.5) +
    scale_fill_manual(values = palette) +
    theme_bw()
  
  if (log_scale){
    plot <- plot + scale_x_log10(breaks = log_breaks, labels = scales::comma)
  }
  
  if (drop_tick_labs){
    plot <- plot + 
              rremove("xlab") +
              rremove("ylab") +
              theme(axis.ticks.x = element_blank(),  
                        axis.text.x = element_blank()) 
  }
  return(plot)
} 
```




# cGVHD time discretization

```{r}

label_data <- data.frame(
  x = c((90 + 120)/2, (120 + 210)/2, (210 + 365)/2, (365 + 500)/2),
  y = 0.0073,
  label = c("Early\ncGVHD", "Intermediate\ncGVHD", 
            "Late\ncGVHD", "Extra Late\ncGVHD")
)

data %>% 
  ggplot() +
  geom_density(aes(x = cGVHD_time), 
               fill = "#1A3D6B", alpha = 0.7) +
  geom_vline(xintercept = c(120, 210, 365), color = "firebrick") +
  geom_label(data = label_data, aes(x = x, y = y, label = label), 
             vjust = -0.5, 
             fill = "white",
             position = position_dodge(width = 0.9), )+
  ylim(c(0, 0.0085)) +
  theme_bw()
```

```{r}
data <- data %>% 
  mutate(cGVHD_time_discrete = factor(case_when(
    cGVHD_time < 120 ~ "early", 
    cGVHD_time >= 120 & cGVHD_time < 210 ~ "intermediate",
    cGVHD_time >= 210 & cGVHD_time < 365 ~ "late",
    cGVHD_time >= 365 ~ "extralate"
  ), levels = c("early", "intermediate", "late", "extralate", NA)))
```

```{r}

```



# CD4+ compartment evaluation
## Major CD4+ T helper populations

```{r}
density <- data %>% 
  select(starts_with("4_"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4_"), names_to = "Population", values_to = "Count") %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4_"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4_"), names_to = "Population", values_to = "Count") %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4_"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4_"), names_to = "Population", values_to = "Count") %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "early" = "#3D527A", 
  "intermediate" = "#717788", 
  "late" = "#AF737D", 
  "extralate" = "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4_"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4_"), names_to = "Population", values_to = "Count") %>%  
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=8, fig.width=15}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, 
          align = "v")
```

```{r}
data %>% 
  select(starts_with("4_")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```

## CD4+ compartment by specific markers

```{r}
density <- data %>% 
  select(starts_with("4+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4+"), names_to = "Population", values_to = "Count") %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4+"), names_to = "Population", values_to = "Count") %>%  
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
  
```

## CD4+ Naïve T cells

```{r}
density <- data %>% 
  select(starts_with("4NV") & !starts_with("4NV+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4NV") & !starts_with("4NV+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4NV") & !starts_with("4NV+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4NV") & !starts_with("4NV+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4NV") & !starts_with("4NV+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4NV") & !starts_with("4NV+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4NV") & !starts_with("4NV+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4NV") & !starts_with("4NV+"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4NV") & !starts_with("4NV+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```

## CD4+ Naïve T cells by specific markers

```{r}
density <- data %>% 
  select(starts_with("4NV+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4NV+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4NV+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4NV+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4NV+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4NV+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4NV+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4NV+"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4NV+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```

## CD4+ Effector Memory T cells

```{r}
density <- data %>% 
  select(starts_with("4EM") & !starts_with("4EM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4EM") & !starts_with("4EM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4EM") & !starts_with("4EM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4EM") & !starts_with("4EM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4EM") & !starts_with("4EM+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4EM") & !starts_with("4EM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4EM") & !starts_with("4EM+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4EM") & !starts_with("4EM+"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4EM") & !starts_with("4EM+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```

## CD4+ Effector Memory T cells by specific markers

```{r}
density <- data %>% 
  select(starts_with("4EM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4EM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4EM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4EM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4EM+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4EM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4EM+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4EM+"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4EM+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```

## CD4+ Central Memory T cells

```{r}
density <- data %>% 
  select(starts_with("4CM") & !starts_with("4CM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4CM") & !starts_with("4CM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4CM") & !starts_with("4CM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4CM") & !starts_with("4CM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4CM") & !starts_with("4CM+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4CM") & !starts_with("4CM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4CM") & !starts_with("4CM+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4CM") & !starts_with("4CM+"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4CM") & !starts_with("4CM+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
  
```

## CD4+ Effector Memory T cells by specific markers

```{r}
density <- data %>% 
  select(starts_with("4CM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4CM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4CM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4CM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4CM+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4CM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4CM+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4CM+"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4CM+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```

## CD4+ Memory T cell compartment

```{r}
density <- data %>% 
  select(starts_with("4TM") & !starts_with("4TM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4TM") & !starts_with("4TM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4TM") & !starts_with("4TM+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4TM") & !starts_with("4TM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4TM") & !starts_with("4TM+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4TM") & !starts_with("4TM+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4TM") & !starts_with("4TM+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4TM") & !starts_with("4TM+"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4TM") & !starts_with("4TM+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```

## CD4+ Effector T cells compartment

```{r}
density <- data %>% 
  select(starts_with("4TE") & !starts_with("4TE+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4TE"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4TE") & !starts_with("4TE+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4TE") & !starts_with("4TE+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4TE") & !starts_with("4TE+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4TE"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4TE") & !starts_with("4TE+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4TE"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4TE") & !starts_with("4TE+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
  
```

## CD4+ Effector T cells compartment by specific markers

```{r}
density <- data %>% 
  select(starts_with("4TE+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("4TE"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("4TE+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(starts_with("4TE+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("4TE+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("4TE"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("4TE+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("4TE"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("4TE+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
  
```

## СD4+ Tregs

```{r}
density <- data %>% 
  select(contains("4TREG"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = contains("TREG"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(contains("4TREG"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(contains("TREG"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(contains("4TREG"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = contains("TREG"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(contains("4TREG"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = contains("TREG"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(contains("4TREG")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
  
```

# CD8+ compartment evaluation

## CD8+ compartment by specific markers

```{r}
density <- data %>% 
  select(starts_with("8+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("8+"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8+"), names_to = "Population", values_to = "Count") %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("8+"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("8+"), names_to = "Population", values_to = "Count") %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("8+"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("8+"), names_to = "Population", values_to = "Count") %>%  
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("8+")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
  
```

## Effector Memory CD8+ T cells

```{r}
density <- data %>% 
  select(starts_with("8EMTM"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8EMTM"), names_to = "Population", values_to = "Count") %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("8EMTM"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8EMTM"), names_to = "Population", values_to = "Count") %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("8EMTM"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("8EMTM"), names_to = "Population", values_to = "Count") %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("8EMTM"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("8EMTM"), names_to = "Population", values_to = "Count") %>%  
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("8EMTM")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
  
```

## Naïve CD8+ T cells

```{r}
density <- data %>% 
  select(starts_with("8NV"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8NV"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("8NV"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8NV"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("8NV"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("8NV"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("8NV"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("8NV"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("8NV")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```

## Central Memory CD8+ T cells

```{r}
density <- data %>% 
  select(starts_with("8CM"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8CM"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("8CM"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8CM"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("8CM"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("8CM"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("8CM"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("8CM"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("8CM")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```

## Effector CD8+ T cells

```{r}
density <- data %>% 
  select(starts_with("8TE"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8TE"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  mutate(Count = Count + 1) %>% 
  ridge_plot(palette = palette, log_scale = TRUE, drop_tick_labs = FALSE)

```

```{r}

time_palette <- c("#21395F", "#3D527A", "#5E729C", "#8093BF", "#A3B6E4")

box_by_day <- data %>% 
  select(starts_with("8TE"), Collection_Day) %>% 
  filter(!is.na(Collection_Day)) %>% 
  pivot_longer(cols = starts_with("8TE"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(Collection_Day = as.factor(Collection_Day)) %>% 
  ggplot(aes(x = Population, y = Count, fill = Collection_Day)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}

cgvhd_palette <- c("#AF737D", "#7083AF")

box_by_cgvhd <- data %>% 
  select(starts_with("8TE"), cGVHD_time) %>% 
  mutate(cGVHD_time = if_else(is.na(cGVHD_time), "No cGVHD", "cGVHD")) %>% 
  pivot_longer(cols = starts_with("8TE"), names_to = "Population", values_to = "Count") %>% 
  mutate(Count = Count + 1) %>% 
  mutate(cGVHD_time = as.factor(cGVHD_time)) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r}
cgvhd_time_palette <- c(
  "#3D527A", "#717788", "#AF737D", "#79424D"
)

box_by_cgvhd_time <- data %>% 
  select(starts_with("8TE"), cGVHD_time_discrete) %>% 
  filter(!is.na(cGVHD_time_discrete)) %>% 
  pivot_longer(cols = starts_with("8TE"), names_to = "Population", values_to = "Count") %>%  
  mutate(Count = Count + 1) %>% 
  ggplot(aes(x = Population, y = Count, fill = cGVHD_time_discrete)) +
  geom_boxplot(position='dodge', outliers = F) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), 
              color='black', shape = 21, size = 0.5) +
  scale_fill_manual(values = cgvhd_time_palette) +
  scale_y_log10(breaks = breaks, labels = scales::comma) +
  theme_bw()
```

```{r, fig.hight=5, fig.width=20}
ggarrange(plotlist = list(density, box_by_day, box_by_cgvhd, box_by_cgvhd_time), 
          ncol = 2, nrow = 2, align = "v")
```

```{r, fig.hight=10, fig.width=10}
data %>% 
  select(starts_with("8TE")) %>% 
  log1p() %>% 
  ggpairs(progress = F, lower = list(continuous = wrap("smooth", 
                                                       alpha = 0.3, 
                                                       size = 1,
                                                       color = "#9B3D53")),
          diag = list(continuous = wrap("densityDiag", fill = "#9B3D53", alpha = 0.5))) +
  theme_bw() +
  theme(strip.text = element_text(size = 5))
```











